---
title: "US-AMS AmeriFlux Formatting"
author: "Gavin McNicol"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load libraries

```{r}
library(tidyverse)
library(skimr)
```

Get data

```{r}
data <- read_csv("data/ATMOS_L3_2024-08-12_ch4gapfilled_fluxnet.csv")
```

```{r}
skim(data)
```

## AmeriFlux Fail 1 Are Timestamp variables present?	FAIL 
Expected timestamp variable(s) TIMESTAMP_START is / are missing.

Check what `TIMESTAMP` corresponds with. 

```{r}
data %>% 
  select(contains("TIMESTAMP"))
```

Shift TIMESTAMP down by one row and fill first cell. Then rename as TIMESTAMP_START. Gather TIMESTAMP_END to second row:

```{r}
data_1 <- data %>% 
  mutate(TIMESTAMP = lag(TIMESTAMP, 1, 202205241000)) %>% 
  rename(TIMESTAMP_START = TIMESTAMP) 
```

## AmeriFlux Fail Are Timestamp variables as expected?	FAIL
These unexpected variables were found in columns 1 & 2 instead of TIMESTAMP_START and TIMESTAMP_END: TIMESTAMP, DATE

Move TIMESTAMP_END to second column. Delete `Timestamp` redundant variable.

```{r}
data_2 <- data_1 %>% 
  select(TIMESTAMP_START, TIMESTAMP_END, everything()) %>% 
  select(-Timestamp) 
```


## AmeriFlux Fail Any invalid Missing-Value Formats?	FAIL
Missing values are not indicated with -9999 for these variables (number of timestamps): 
MANY vars.    


Missing values are filled with double type not integer -9999. Try fixing:  

- write function to take vector, convert -9999 to NA (Have to remove date type columns!)
- convert data_2 into a list of vectors
- map function to list
- convert list back to data_2 data_frame
- after writing out data as .csv, replace NA with -9999 integer

```{r}
fix_missing <- function(vector){
  
  if(typeof(vector) == "double"){
    vector %>% 
      na_if(-9999) %>% 
      na_if(9999.99)
  }
  else
  {
    vector
  }
} 
```

test function: works except on date or datetime vars so remove those from data_2

```{r}
x <- as.POSIXct( c(1, 2, 3, -9999, 9999.99) )
typeof(x)
fix_missing(x) 
```

remove all date-and-time type columns

```{r}
data_2 <- data_2 %>% 
  select(-date, -DATE, -datetime, -time, -time_local, -time_local.met)
```

turn data_2 into a list

```{r}
data_2_list <- as.list(data_2)
```

map custom na_if function

```{r}
data_2_list <- map(data_2_list, fix_missing)
```

convert back to data frame

```{r}
data_3 <- as_tibble(data_2_list)
# skim(data_3)
```

## AmeriFlux Fail Issues that cannot be autocorrected.	FAIL
File has duplicate variables DATE (column 4); TIME (column 277); TIMESTAMP (column 432); YEAR (column 434). File had issues that could not be automatically corrected. Autocorrection FAILED.

check to see if these have been completely removed:

`year` is still present

```{r}
data_3 %>% 
  select(year)
```

```{r}
data_3 <- 
  data_3 %>% 
  select(-year)
```


# File name fixes

## AmeriFlux Fail Timestamp problem encountered.	FAIL
These Format QA/QC assessments could not be completed: Do filename time components match file time period? Is Timestamp resolution OK? Any Timestamp duplicates?

## AmeriFlux Fail Is Filename Format valid?	FAIL
Is Filename Format valid?	FAIL

## AmeriFlux Fail Is Filename Format valid?	WARNING
These filename components are not in the standard AmeriFlux format: ts-start (start time), optional parameter included (will be removed in autocorrected file)

write out correctly

```{r}
names(data_3)
```


```{r}
data_3 %>% 
  write_csv("data/US-AMS_HH_202205241000_202312201500.csv", na = "-9999")
```


